---
title: "ggplot2 exercises"
author: "Leonard Blaschek"
output: html_document
---

```{r setup, include=FALSE}
# in this "setup chunk" you can load packages:
library(tidyverse)

# you can also create a custom theme_*():
theme_sulis <- function(base_size = 8) {
  theme_minimal(
    base_size = base_size
  ) %+replace%
    theme(
      axis.ticks = element_blank(),
      axis.text.x = element_text(
        size = 8,
        colour = "black",
        margin = margin(1, 1, 1, 1)
      ),
      axis.text.y = element_text(
        colour = "black",
        size = 8,
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      axis.title = element_text(
        colour = "black",
        size = 8
      ),
      axis.line = element_line(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.text = element_text(size = 8),
      legend.key.height = unit(4, "mm"),
      complete = TRUE
    )
}

# and define variables you might need:
ggtext_size <- 8 / (14 / 5)
```

## 1. Density curves

Task: Create density curves (`geom_density()`) for the relative lignin content (`rel_lignin`).
The fill of the curve should correspond to their `type` (mutant or WT) and be semi-transparent (`alpha = 0.5`). 
Lastly, get rid of the hideous grey background by applying our custom `theme_sulis()`.

Data file: `"data/Sulis2023_fig3G.tsv"`

```{r}
# write your code in this code chunk; execute all code within the chunk your cursor is in with ctrl+shift+enter
data <- read_tsv("data/Sulis2023_fig3G.tsv")

ggplot(data = data,
       aes(
         x = rel_lignin,
         fill = type)) +
  geom_density(alpha = 0.5) +
  theme_sulis()
```

## 2. Violin and boxplots

Task: Plot the distribution of predicted carbohydrate/lignin ratios (`rel_CL`) as a boxplots (`geom_boxplot()`) on top of violin plots (`geom_violin()`), grouped by the number mutated genes (`target_genes`).
Reduce the width of the boxplots to fit within the violin plots (check `?geom_boxplot()` for the right argument).
Fill the violin plots according to the number of mutated genes, using the continuous viridis colour scale (`scale_colour_viridis_c()`).

Data: `"data/Sulis2023_predictions.tsv"`

Hint: If you want to use a continuous variable as a grouping factor, you might have to wrap it inside `as.factor()`.

```{r}
data <- read_tsv("data/Sulis2023_predictions.tsv")
ggplot(data = data,
       aes(x = as.factor(target_genes),
           y = rel_CL)) +
  geom_violin(aes(fill = target_genes)) +
  geom_boxplot(width = 0.05) +
  scale_fill_viridis_c() +
  theme_sulis()
```


