---
title: "ggplot2 exercises"
author: "Leonard Blaschek"
output: html_document
---

```{r setup, include=FALSE}
# in this "setup chunk" you can load packages:
library(tidyverse)

# you can also create a custom theme_*():
theme_sulis <- function(base_size = 8) {
  theme_minimal(
    base_size = base_size
  ) %+replace%
    theme(
      axis.ticks = element_blank(),
      axis.text.x = element_text(
        size = base_size,
        colour = "black",
        margin = margin(1, 1, 1, 1)
      ),
      axis.text.y = element_text(
        colour = "black",
        size = base_size,
        angle = 0,
        vjust = 0.5,
        hjust = 1,
        margin = margin(1, 1, 1, 1)
      ),
      axis.title = element_text(
        colour = "black",
        size = base_size
      ),
      axis.line = element_line(),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.text = element_text(size = base_size),
      legend.key.height = unit(4, "mm"),
      complete = TRUE
    )
}

# and define variables you might need:
ggtext_size <- 8 / (14 / 5)
```

## 1. Density curves

#### Task: 
Create density curves (`geom_density()`) for the relative lignin content (`rel_lignin`).
The fill of the curve should correspond to their `type` (mutant or WT) and be semi-transparent (`alpha = 0.5`). 
Lastly, get rid of the hideous grey background by applying our custom `theme_sulis()`. 

#### Bonus: 
Check the arguments of `coord_cartesian()` to figure out how to remove the scale expansion (*i.e.* the little gap between 0 and the axis line).

#### Data file: 
`"data/Sulis2023_fig3G.tsv"`

```{r}
# write your code in this code chunk; execute all code within the chunk your cursor is in with ctrl+shift+enter
data <- read_tsv("data/Sulis2023_fig3G.tsv")

ggplot(
  data = data,
  aes(
    x = rel_lignin,
    fill = type
  )
) +
  geom_density(alpha = 0.5) +
  coord_cartesian(expand = FALSE) +
  theme_sulis()
```

## 2. Violin and boxplots

#### Task: 
Plot the distribution of predicted carbohydrate/lignin ratios (`rel_CL`) as boxplots (`geom_boxplot()`) on top of violin plots (`geom_violin()`), grouped by the number mutated genes (`target_genes`).
Reduce the width of the boxplots to fit within the violin plots (check `?geom_boxplot()` for the right argument).
Fill the violin plots according to the number of mutated genes, using the continuous viridis colour scale (`scale_colour_viridis_c()`), which is designed to be interpretable by viewers with common forms of colour blindness.

#### Data: 
`"data/Sulis2023_predictions.tsv"`

#### Hint: 
If you want to use a continuous variable as a grouping factor on a discrete x-axis, you might have to wrap it inside `as.factor()`.

```{r}
data <- read_tsv("data/Sulis2023_predictions.tsv")
ggplot(
  data = data,
  aes(
    x = as.factor(target_genes),
    y = rel_CL
  )
) +
  geom_violin(aes(fill = target_genes)) +
  geom_boxplot(width = 0.05) +
  scale_fill_viridis_c() +
  theme_sulis()
```

# 3. 2D density and facets

#### Task: 
Using the data of mutant predictions, plot the relative stem volume (`rel_volume`) against relative lignin content (`rel_lignin`).
You can start by trying `geom_point()`, but you'll see that the enormous number of points creates a black mass that is difficult to interpret.
Instead, try `geom_density2d_filled()` to get a 2D density map of the predictions.

This looks OK, but what if you want to compare those densities relative to the number of targeted genes?
`facet_wrap()` lets you split your figure into smaller facets, each representing a subgroup of the data.
Use `facet_wrap()` to split your plot by `target_genes`.
Lastly, add white dashed lines at 100% of x and y axes for orientation (check `?geom_abline()`).

#### Bonus:
Seven facets look a bit inelegant, and the `0` facet (*i.e.* WT) is useless anyways.
Going a bit beyond the scope of this workshop, replace your data argument with
`data = data |> filter(target_genes > 0)` to remove the `0` facet (assuming you named your data frame `data`).

#### Data: 
`"data/Sulis2023_predictions.tsv"`

#### Hint: 
The unequal number of data points in the different facets makes the densities in facets with fewer target genes invisible.
To calculate the bins independently in each facet, set the `geom_density2d_filled()` option `contour_var` to `"ndensity"`.

```{r}
data <- read_tsv("data/Sulis2023_predictions.tsv")
ggplot(
  data = data |> filter(target_genes > 0),
  aes(
    x = rel_lignin,
    y = rel_volume
  )
) +
  geom_density2d_filled(contour_var = "ndensity") +
  geom_hline(yintercept = 100, colour = "white", linetype = "dashed") +
  geom_vline(xintercept = 100, colour = "white", linetype = "dashed") +
  theme_sulis() +
  facet_wrap(~target_genes)
```
